Music Recommendation System Using the Million Song Dataset

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex
Updating .tex
Removed "in reverse" when describing the MAP@500 because it's not true.
Added a sentence about sparsity in PMF. Still figuring out how to talk
about running time and space requirements for PMF

Intuitively, the score looks at the percentage of recommendations that were in-line with what the user actually liked using the testing set that were not used by the algorithms during training. This score also takes into account the order of the songs recommended so it is desirable to have the song with the highest confidence to be liked at the beginning of the recommendation list. For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 1 while recommending them at the very end of the 500 songs in reverse order would only produce a MAP score of 0.0060. The MAP score can be said to be similar to the AUC of the ROC curve discussed in class without taking into account the order of recommendations.

@@

Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline. Based on literature review \citep{li2012million, McFee:2012:MSD:2187980.2188222}, we have found that this is better or in line with what others have achieved with using PMF on play counts. Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse. However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.

@@

\section{Results}
After comparing the various algorithms, we found that item-based collaborative filtering gave us the best result. Our most complicated model, matrix factorization, gave us results barely above the popularity baseline. On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system. Our song-based collaborative filtering technique produced the best results despite it being more naive as compared to matrix factorization. Our final results for each technique employed are summarized in Table \ref{tab:map}.

>>>>

Intuitively, the score looks at the percentage of recommendations that were in-line with what the user actually liked using the testing set that were not used by the algorithms during training. This score also takes into account the order of the songs recommended so it is desirable to have the song with the highest confidence to be liked at the beginning of the recommendation list. For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 1 while recommending them at the very end of the 500 songs would only produce a MAP score of 0.0060. The MAP score can be said to be similar to the AUC of the ROC curve discussed in class without taking into account the order of recommendations.

@@

Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline. Based on literature review \citep{li2012million, McFee:2012:MSD:2187980.2188222}, we have found that this is better or in line with what others have achieved with using PMF on play counts. Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse. A dataset that is too sparse prevents PMF from learning enough information about how users rank songs to produce more accurate recommendations. However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.

@@

\section{Results}
After comparing the various algorithms, we found that item-based collaborative filtering gave us the best result. Our most complicated model, matrix factorization, gave us results barely above the popularity baseline. On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system. This is probably because users tend to listen to multiple songs by the same artist. Our song-based collaborative filtering technique produced the best results despite it being more naive as compared to matrix factorization. Our final results for each technique employed are summarized in Table \ref{tab:map}.
